\chapter{Recurrent reorganization of the three-dimensional genome pinpoint non-coding drivers of primary prostate tumours}
\chaptermark{CRE Hijacking and the 3D Prostate Genome}
\label{chap:3D}

This chapter is a version of the preprint published on bioRxiv and currently under review:

\textbf{Hawley, J. R.}*, Zhou, S.*, Arlidge, C., Grillo, G., Kron, K. J., Hugh-White, R., van der Kwast, T., Fraser, M., Boutros, P. C., Bristow, R. G., and Lupien, M.
\underline{\emph{Cis}-regulatory Element Hijacking by} \underline{Structural Variants Overshadows Higher-Order Topological Changes in Prostate Cancer}.
\textbf{bioRxiv}, 2021.
doi: 10.1101/2021.01.05.425333.

* The authors co-led the study with equal contributions and can be interchangeably listed as first author.

\vspace{1em}

Contributions per the manuscript:
S.Z., J.R.H., and M.L. conceptualized the study.
J.R.H. and S.Z. co-led the study with equal contributions and can be interchangeably listed as first author.
S.Z. designed and conducted all the experiments with help from C.A. G.G., and K.K.
J.R.H. implemented all the computational and statistical approaches and analyses.
R.H.-W. pre-processed the \gls{rnaseq} data from the primary tumours.
Figures were designed by S.Z. and J.R.H.
The manuscript was written by S.Z., J.H., and M.L with assistance from all authors.
T.v.d.K., M.F., P.C.B., R.G.B., and M.L. supervised the study.
M.L. oversaw the study.

\subfile{chapter3/abstract}
\subfile{chapter3/intro}
\subfile{chapter3/results}
\subfile{chapter3/discussion}
\subfile{chapter3/methods}
