\section*{List of Research Works}

\subsection*{Full Manuscripts}

\subsubsection*{Published}

\begin{itemize}
  \item \fullcite{seneviratneMitochondrialTransacylaseTafazzin2019}
  \item \fullcite{singhDisruptingMitochondrialCopper2020}
  \item \fullcite{zhouNoncodingMutationsTarget2020}
\end{itemize}


\subsubsection*{In Preparation or Submitted}

\begin{itemize}
  \item \fullcite{hawleyRecurrentReorganizationThreedimensional2021}
  \item \fullcite{ghamrasniNoncodingMutationsReveal2021}
\end{itemize}

\subsection*{Selected talks and first author poster presentations}

\begin{itemize}
  \item "\Glsentrylong{cre} hijacking overshadows higher-order topological changes in \glsentrylong{pca}". Fragile Nucleosome Seminar. Remote, 2021.
  \item "\Glsentrylong{cre} hijacking overshadows topological changes in \glsentrylong{pca}". Toronto Bioinformatics Users Group (TorBUG) Seminar. Toronto, Ontario, 2021.
  \item "\Glsentrylong{cre} hijacking overshadows topological changes in \glsentrylong{pca}". Toronto Transcription, Chromatin, and Evolution Seminar. Toronto, Ontario, 2021.
  \item "Heterogeneity of \glsentrylong{dname} \glsentrylong{ball} relapse". Keystone Symposium - Epigenetics and Human Disease. Banff, British Columbia, 2019.
  \item "Epigenetic dynamics in relapse of \glsentrylong{ball}". Canadian Epigenetics, Environment, and Health Research Consortium Conference. Banff, British Columbia, 2019.
  \item "\glsentrylong{dname} heterogeneity of \glsentrylong{ball} and cancer cell lines". Ontario Institute for Cancer Research Leukemia Translational Research Initiative Symposium. Toronto, Ontario, 2018.
  \item "Assessing epigenomic heterogeneity through DNA methylation entropy in cancer cells". Canadian Epigenetics, Environment, and Health Research Consortium Conference. Esterel, Quebec, 2018.
  \item "Assessing epigenetic heterogeneity through DNA methylation entropy in cancer cell lines". James Lepock Memorial Symposium. Toronto, Ontario, 2018.
\end{itemize}
